Learning to Infer Graphics Programs from Hand-Drawn Images

### 2017-10-25 short.tex
started work on shorter version of the paper: see short.tex

 How can an agent convert noisy, high-dimensional perceptual input
 to a symbolic, abstract object, such as a computer program?  Here we
 consider this problem within a graphics program synthesis domain.  We
 develop an approach for converting  hand
 drawings into executable source code for drawing the original image.
 The graphics programs in our domain draw simple figures like those found in
 machine learning papers (see Figure~\ref{firstPageExamples}a).
 \begin{figure}[H]\vspace{-0.7cm}
  \begin{minipage}[t]{0.7\linewidth}  
\begin{tabular}{llll}
  \includegraphics[width = \exampleImageSize]{figures/expert-60.png}&
  \includegraphics[width = \exampleImageSize]{figures/expert-5.png}&
    \includegraphics[width = \exampleImageSize]{figures/expert-17.png}&
    \includegraphics[width = \exampleImageSize]{figures/expert-58.png}\\
  \includegraphics[width = \exampleImageSize]{figures/60.png}&
  \includegraphics[width = \exampleImageSize]{figures/5.png}&
    \includegraphics[width = \exampleImageSize]{figures/17.png}&
    \includegraphics[width = \exampleImageSize]{figures/58.png}
\end{tabular}
\subcaption{}
  \end{minipage}%
  \begin{minipage}[t]{0.3\linewidth}
    \begin{tikzpicture}
      \node(picture) at (0,0) {\includegraphics[width = 2cm]{figures/expert-31.png}};
      \draw[very thick,->] (picture.south)  -- (0,-2);
    \end{tikzpicture}\\
    \begin{minipage}[t]{0.3\linewidth}
\begin{Verbatim}
for (i < 3)
 rectangle(3*i,-2*i+4,
           3*i+2,6)
 for (j < i + 1)
  circle(3*i+1,-2*j+5)
\end{Verbatim}      
      \end{minipage}
    \subcaption{}
  \end{minipage}
  \caption{(a): Model learns to convert hand drawings (top) into \LaTeX~(rendered below). (b) Synthesizes high-level \emph{graphics program} from hand drawing.}\label{firstPageExamples}
\end{figure}

The key observation behind our work is that generating a programmatic representation from an image of a diagram actually involves two distinct steps that require different technical approaches. The first step involves identifying the components such as rectangles, lines and arrows that make up the image. The second step involves identifying the high-level structure in how the components were drawn. In Figure 1(b), it means identifying a pattern in how the circles and rectangles are being drawn that is best described with two nested loops, and which can easily be extrapolated to a bigger diagram.

We present a hybrid architecture for inferring graphics programs that is structured around these two steps. For the first step, our approach uses a deep network to infer a set of primitive shape-drawing commands. We refer to this set as a \emph{trace},\footnote{For our domain we do not require the trace to be ordered -- so it is similar to, but not exactly the same as, the execution trace of the program.} since it corresponds to the commands that a program would have issued but lacks the high-level structure that determines how the program decided to issue them.
For added robustness, we train the network to produce a proposal distribution of the most likely traces and use stochastic search to find the trace that best matches the image.
%The network is trained from an automatically-generated corpus of synthetic image-generating programs.

The second step involves generating a high-level program capable of producing the trace identified by the first phase.
%In principle, we could also train a neural network to learn this mapping. In practice, though, this would be challenging because it would no longer be sufficient to train the network with a synthetic corpus of randomly generated programs, since our goal is to generate programs that satisfy certain semantic constraints that ensure that we get the program that the user most likely intended.
%This paper argues that this stage is best
We achieve this by \emph{constraint-based program synthesis} \cite{solar2008program}. The program synthesizer can search the space of possible programs for one capable of producing the desired trace -- inducing structures like
symmetries, repetition, or conditional branches.
Although these program do not need any training data,
we show how to learn a \emph{search policy} 
in order to synthesize programs an order of magnitude faster than constraint-based synthesis techniques alone.

@@

The roadmap of our paper is structured around the trace hypothesis, as outlined in Figure~\ref{roadmap}.
\begin{figure}[h]
  \begin{tikzpicture}
    \node[ thick,anchor = west,inner sep=0pt,label=below:{\small Perceptual input}](observation) at (0,0) {\includegraphics[width = 1.5cm]{figures/expert-39-trimmed.png}};
    \node[ultra thick,anchor = west,inner sep=0pt](traceSource) at (3,0.5){    \begin{lstlisting}[basicstyle = \scriptsize\ttfamily]
line, line,
rectangle,
line, ...
\end{lstlisting}};
    \node[ultra thick,anchor = west,inner sep=0pt](traceImage) at (3.22,-0.5) {
      \includegraphics[width = 0.9cm]{figures/39-parse.png}}; 
    \node(trace)[draw,thin,fit = (traceImage) (traceSource), label = below:{{\small Trace}}] {};

@@

    \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
    ([yshift = -15,xshift = -115]trace.south) -- node[below = 6pt] {Section~\ref{neuralNetworkSection}: Perception$\to$Trace} ([yshift = -15,xshift = -5]trace.south);
    \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
    ([xshift = 5,yshift = -15]trace.south) -- node[below = 6pt] {Section~\ref{programSynthesisSection}: Trace$\to$Program} ([xshift = 140,yshift = -15]trace.south);
    \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
([xshift = 150,yshift = -15]trace.south) -- node[below = 6pt] {Section~\ref{applicationsSection}: Applications} ([xshift = 250,yshift = -15]trace.south);
  \end{tikzpicture}
  \caption{Paper roadmap. Both the paper and the system pipeline are structured around the \emph{trace hypothesis}}\label{roadmap}
\end{figure}

Prior work addresses converting hand drawings to vectors~\cite{huang2017shape,Nishida:2016:ISU:2897824.2925951} and converting synthetic images to symbolic parses~\cite{nsd,DBLP:journals/corr/Beltramelli17}.
Our new contributions are:
\begin{itemize}
\item [-] A model that converts sketches to high-level programs
\item [-] An algorithm for learning a policy for efficiently searching for programs
\item [-] The trace hypothesis: a framework for going from perception to programs
  \end{itemize}

@@

%% The key observation behind our work is that generating a programmatic representation from an image of a diagram actually involves two distinct steps that require different technical approaches. The first step involves identifying the components such as boxes, lines and arrows that make up the image. The second step involves identifying the high-level structure in how the components were drawn; for example, in Figure 1(b), it means identifying that there is a pattern in how the circles and rectangles are being drawn that is best described with two nested loops, and which can easily be extrapolated to a bigger diagram.

@@

Figure~\ref{architecture} illustrates this architecture.  We first
pass a $256\times 256$ target image and a rendering of the trace so
far (encoded as a two-channel image) to a convolutional network. Given
the features extracted by the convnet, a multilayer perceptron then
predicts a distribution over the next drawing command to add to the
trace; see Table~\ref{drawingCommandTable} for the drawing commands we currently model.  We predict the drawing
command token-by-token,\footnote{A token is an atomic symbol  -- so the model is \emph{not} predicting lines of code character-by-character, but instead knows about the syntactic structure of primitive commands} conditioning each token both on the image
features and on the previously generated tokens. We also use a
differentiable attention mechanism (Spatial Transformer
Networks:~\cite{jaderberg2015spatial}) to let the model attend to
different regions of the image while predicting each token. For
example, the network first decides to emit the \verb|circle| token
conditioned on the image features, then it emits the $x$ coordinate of
the circle conditioned on the image features and the \verb|circle|
token, and finally it predicts the $y$ coordinate of the circle
conditioned on the image features, the \verb|circle| token, and the
$x$ coordinate. For our main experiments we constrain
coordinates to lie on a discrete $16\times 16$ grid, and later explain
how to remove this restriction.  See supplement for the full details
of the architecture, which we implemented in
Tensorflow~\cite{tensorflow2015-whitepaper}.
%% \remark{There are some more details that are important to provide
%%   about this architecture in the supplement: the
%%   functional form(s) of the probability distributions over tokens, the
%%   network layer sizes, which MLPs share parameters, etc.}

@@

Training does not require backpropagation across the entire sequence
of drawing commands: drawing to the canvas `blocks' the gradients,
effectively offloading memory to an external visual store.  In a
sense, this model is like an autoregressive variant of
AIR~\cite{eslami1603attend}, but critically, \emph{without a learned
  recurrent memory state}, e.g. an RNN.  We can do without an RNN
because we have access to ground truth (image, trace) pairs.  This
allows handling scenes with a very large number of objects without
having to worry about the conditioning of gradients as they propagate
over long sequences, and also makes training more straightforward (it
is just maximum likelihood estimation of the model parameters).

Our network suffices to ``derender'' synthetic images like those shown in
Figure~\ref{trainingData}.  We can perform a beam search decoding to
recover what the network thinks is the most likely trace for
images like these, recovering traces maximizing $\probability_\theta
[T|I]$. But, if the network makes a mistake (predicts an incorrect
line of code), it has no way of recovering from the error.  In order
to derender an image with $n$ objects, it must correctly predict $n$
drawing commands -- so its probability of success will decrease
exponentially in $n$, assuming it has any nonzero chance of making a
mistake.  For added robustness as $n$ becomes large, we treat the
neural network outputs as proposals for a Sequential Monte Carlo (SMC) sampling scheme~\cite{SMCBook}.  For
the SMC sampler, we use pixel-wise distance as a surrogate for a
likelihood function. The SMC sampler is designed to produce samples
from the distribution $\propto L(I|\text{render}(T))
\probability_\theta[T|I]$, where $L(\cdot | \cdot):\text{image}^2\to
\mathbb{R}$ uses the distance between two images as a proxy for a
likelihood. Unconventionally, the target distribution of the SMC sampler
includes the likelihood under the proposal distribution.
Intuitively, both the proposal
distribution and the distance function offer complementary signals for
whether a drawing command is correct. Empirically, accuracy is much
better when we combine these signals.

Figure~\ref{syntheticResults}
compares the neural network with SMC against the neural network by
itself or SMC by itself.  Only the combination of the two passes a
critical test of generalization: when trained on images with $\leq 12$
objects, it successfully parses scenes with many more objects than the
training data.
We also compare with a natural baseline that models the problem as a kind of image captioning (``LSTM'' in Figure~\ref{syntheticResults}).
Given the target image, this baseline produces the program trace in one shot by
using a CNN to extract features of the input which are passed to an LSTM which finally predicts
the trace token-by-token.
This architecture is used in several successful neural models of image captioning (e.g.,~\cite{vinyals2015show}),
but, for this domain, does not suffice to parse cluttered scenes with many objects.

@@

\subsection{Generalizing to hand drawings}
A practical application of our neural network is the automatic conversion of hand drawings into a subset of \LaTeX.
 We train the model
to generalize to hand drawings by introducing noise into the
renderings of the training target images.
We designed this noise process to introduce the kinds of variations found in hand drawings (Figure~\ref{handDrawingExamples}; see supplement for details).

@@

Our neurally-guided SMC procedure
used pixel-wise distance as a surrogate for a likelihood function ($L(\cdot|\cdot)$ in section~\ref{neuralNetworkSection}).
But pixel-wise distance fares poorly on hand drawings, which never exactly match
the model's renders.
So, for hand drawings,
we \emph{learn} a surrogate likelihood function,
$L_{\text{learned}}(\cdot|\cdot)$.
The density $L_{\text{learned}}(\cdot|\cdot)$ is predicted by a convolutional network that we train to predict
the distance between two traces conditioned upon their renderings.
We train our likelihood surrogate to approximate the symmetric difference,
which is  the number of drawing commands by which two traces differ:
\begin{multline}
-\log L_{\text{learned}}(\text{render}(T_1)|\text{render}(T_2))\approx\\ |T_1 - T_2| + |T_2 - T_1|\label{symmetricDistance}
\end{multline}
Intuitively,
$L_{\text{learned}}(\cdot|\cdot)$ approximates the distance between
the trace we want and the trace we have so far.  Pixel-wise distance
metrics are sensitive to the details of how
arrows, dashes, and corners are drawn -- but we wish to be invariant
to these details. So, we learn a distance metric over images that
approximates the distance metric in the search space over traces.

We drew 100 figures by hand; see figure~\ref{lotsOfHandDrawings}.
These were drawn carefully but not perfectly.
Because our model assumes that objects are snapped to a $16\times 16$ grid, 
we made the drawings on graph paper.

For each drawing we annotated a ground truth trace and asked the neurally guided SMC sampler
to produce many candidate traces for each drawing.
For 63\% of the drawings, the Top-1 most likely sample exactly matches the
ground truth; with more samples, the model finds traces
that are closer to the ground truth annotation (Fig.~\ref{drawingIntersectionOverUnion}).
Because our current model sometimes makes mistakes on hand drawings,
we envision it working as follows:
a user sketches a diagram,
and the system responds by proposing a few candidate interpretations.
The user could then select the one closest to their intention and edit it if necessary.

@@

We constrain the space of allowed programs by writing down a context
free grammar over a space of programs. Although it might be desirable
to synthesize programs in a Turing-complete language such as Lisp or
Python, a more tractable approach is to specify what in the program
languages community is called a Domain Specific Language (DSL)~\cite{polozov2015flashmeta}. Our DSL (Table~\ref{DSL})
encodes prior knowledge of what graphics programs tend to look like.

@@

Given the DSL and a trace $T$, we want to recover a program that both evaluates to $T$
and, at the same time, is the ``best'' explanation of $T$.
For example, we might prefer more general programs or, in the spirit of Occam's razor,
prefer shorter programs.
We wrap these intuitions up into a cost function over programs,
and seek the minimum cost program consistent with $T$:
\begin{equation}
  \text{program}(T) = \argmin_{p\in \text{DSL, s.t. }p \text{ evaluates to } T} \text{cost}(p)\label{programObjective}
\end{equation}
We define the
cost of a program to be the number of Statement's it contains (Table~\ref{DSL}).
We also penalize using many different numerical constants; see supplement.
%\remark{The flow here is a bit off/backwards. ``We want a program that evaluates to $T$ and also minimizes some measure of program cost''---why do we care about cost? It'd be better to start by making a ``Bayesian Occam's razor'' appeal (e.g. the most compact/general program is the more likely explanation) and then saying that one way to do this is to minimize a cost function which is proportional to program length.}

The constrained optimization problem in
Equation~\ref{programObjective} is intractable in general, but there
exist efficient-in-practice tools for finding exact solutions to such
program synthesis problems. We use the state-of-the-art Sketch
tool~\cite{solar2008program}.
%Describing Sketch's program synthesis
%algorithm is beyond the scope of this paper; see~\cite{solar2008program}.
At a
high level, Sketch takes as input a space of programs, along with a
specification of the program's behavior and optionally a cost
function.  It translates the synthesis problem into a constraint
satisfaction problem and then uses a SAT solver to find a
minimum-cost program satisfying the specification.  In exchange for
having no guarantee on the time required to find a minimum
cost solution, it comes with the guarantee that it will always find a
globally optimal program.
In practice, synthesis times vary from minutes to hours,
with 16\% of the diagrams timing out the synthesizer after 3 hours.
See Figure~\ref{exampleSynthesisResults} for examples of the kinds of programs our system recovers.
A main impediment to our use of these general techniques is
the prohibitively high cost of searching for programs.
We next describe how to learn to synthesize programs much faster (Section~\ref{learningASearchPolicy}),
timing out on 2\% of the drawings and solving most problems in under a minute.

@@

\section{Related work}

>>>>

 How can an agent convert noisy, high-dimensional perceptual input
 to a symbolic, abstract object, such as a computer program?  Here we
 consider this problem within a graphics program synthesis domain.  We
 develop an approach for converting  hand
 drawings into executable source code for drawing the original image.
 The graphics programs in our domain draw simple figures like those found in
 machine learning papers (see Fig.~\ref{firstPageExamples}a).
 \begin{figure}[H]\vspace{-0.7cm}
  \begin{minipage}[t]{0.7\linewidth}  
\begin{tabular}{llll}
  \includegraphics[width = \exampleImageSize]{figures/expert-60.png}&
  \includegraphics[width = \exampleImageSize]{figures/expert-5.png}&
    \includegraphics[width = \exampleImageSize]{figures/expert-17.png}&
    \includegraphics[width = \exampleImageSize]{figures/expert-58.png}\\
  \includegraphics[width = \exampleImageSize]{figures/60.png}&
  \includegraphics[width = \exampleImageSize]{figures/5.png}&
    \includegraphics[width = \exampleImageSize]{figures/17.png}&
    \includegraphics[width = \exampleImageSize]{figures/58.png}
\end{tabular}
\subcaption{}
  \end{minipage}%
  \begin{minipage}[t]{0.3\linewidth}
    \begin{tikzpicture}
      \node(picture) at (0,0) {\includegraphics[width = 2cm]{figures/expert-31.png}};
      \draw[very thick,->] (picture.south)  -- (0,-2);
    \end{tikzpicture}\\
    \begin{minipage}[t]{0.3\linewidth}
\begin{Verbatim}
for (i < 3)
 rectangle(3*i,-2*i+4,
           3*i+2,6)
 for (j < i + 1)
  circle(3*i+1,-2*j+5)
\end{Verbatim}      
      \end{minipage}
    \subcaption{}
  \end{minipage}
  \caption{(a): Model learns to convert hand drawings (top) into \LaTeX~(rendered below). (b) Synthesizes high-level \emph{graphics program} from hand drawing.}\label{firstPageExamples}
\end{figure}

The key observation behind our work is that generating a programmatic representation from an image of a diagram actually involves two distinct steps that require different technical approaches. The first step involves identifying the components such as rectangles, lines and arrows that make up the image. The second step involves identifying the high-level structure in how the components were drawn. In Fig. 1(b), it means identifying a pattern in how the circles and rectangles are being drawn that is best described with two nested loops, and which can easily be extrapolated to a bigger diagram.

We present a hybrid architecture for inferring graphics programs that is structured around these two steps. For the first step, our approach uses a deep network to infer a set of primitive shape-drawing commands. We refer to this set as an unordered \emph{trace},\footnote{For our domain we do not require the trace to be ordered -- so it is similar to, but not exactly the same as, the execution trace of the program.} since it corresponds to the commands that a program would have issued but lacks the high-level structure that determines how the program decided to issue them.
%For added robustness, we train the network to produce a proposal distribution of the most likely traces and use stochastic search to find the trace that best matches the image.
%The network is trained from an automatically-generated corpus of synthetic image-generating programs.
The second step involves synthesizing a high-level program capable of producing the trace identified by the first phase.
%In principle, we could also train a neural network to learn this mapping. In practice, though, this would be challenging because it would no longer be sufficient to train the network with a synthetic corpus of randomly generated programs, since our goal is to generate programs that satisfy certain semantic constraints that ensure that we get the program that the user most likely intended.
%This paper argues that this stage is best
We achieve this by \emph{constraint-based program synthesis} \cite{solar2008program}. The program synthesizer can search the space of possible programs for one capable of producing the desired trace -- inducing structures like
symmetries, loops, or conditional branches.
Although these program do not need any training data,
we show how to learn a \emph{search policy} 
in order to synthesize programs an order of magnitude faster than constraint-based synthesis techniques alone.

@@

The roadmap of our paper is structured around the trace hypothesis, as outlined in Fig.~\ref{roadmap}.
\begin{figure}[h]
  \begin{tikzpicture}
    \node[ thick,anchor = west,inner sep=0pt,label=below:{\small Perceptual input}](observation) at (0,0) {\includegraphics[width = 1.5cm]{figures/expert-39-trimmed.png}};
    \node[ultra thick,anchor = west,inner sep=0pt](traceSource) at (3,0.5){    \begin{lstlisting}[basicstyle = \scriptsize\ttfamily]
line, line,
rectangle,
line, ...
\end{lstlisting}};
    \node[ultra thick,anchor = west,inner sep=0pt](traceImage) at (3.22,-0.5) {
      \includegraphics[width = 0.9cm]{figures/39-parse.png}}; 
    \node(trace)[draw,thin,fit = (traceImage) (traceSource), label = below:{{\small Trace}}] {};

@@

    \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
    ([yshift = -15,xshift = -115]trace.south) -- node[below = 6pt] {Section~\ref{neuralNetworkSection}: Perception$\to$Trace} ([yshift = -15,xshift = -5]trace.south);
    \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
    ([xshift = 5,yshift = -15]trace.south) -- node[below = 6pt] {Section~\ref{programSynthesisSection}: Trace$\to$Program} ([xshift = 140,yshift = -15]trace.south);
    \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
([xshift = 150,yshift = -15]trace.south) -- node[below = 6pt] {Section~\ref{applicationsSection}: Applications} ([xshift = 250,yshift = -15]trace.south);
  \end{tikzpicture}
  \caption{Both the paper and the system pipeline are structured around the \emph{trace hypothesis}}\label{roadmap}
\end{figure}

Prior work addresses converting hand drawings to vectors~\cite{huang2017shape,Nishida:2016:ISU:2897824.2925951} and converting synthetic images to symbolic parses~\cite{nsd,DBLP:journals/corr/Beltramelli17}.
Our new contributions are: (1) A model that converts sketches to high-level programs. (2) An algorithm for learning a policy for efficiently searching for programs. (3) The trace hypothesis: a framework for going from perception to programs.
  \end{itemize}

@@

%% The key observation behind our work is that generating a programmatic representation from an image of a diagram actually involves two distinct steps that require different technical approaches. The first step involves identifying the components such as boxes, lines and arrows that make up the image. The second step involves identifying the high-level structure in how the components were drawn; for example, in Fig. 1(b), it means identifying that there is a pattern in how the circles and rectangles are being drawn that is best described with two nested loops, and which can easily be extrapolated to a bigger diagram.

@@

Fig.~\ref{architecture} illustrates this architecture.  We first
pass a $256\times 256$ target image and a rendering of the trace so
far (encoded as a two-channel image) to a convolutional network. Given
the features extracted by the convnet, a multilayer perceptron then
predicts a distribution over the next drawing command to add to the
trace; see Tbl.~\ref{drawingCommandTable} for the drawing commands we currently model.  We predict the drawing
command token-by-token,\footnote{A token is an atomic symbol  -- so the model is \emph{not} predicting lines of code character-by-character, but instead knows about the syntactic structure of primitive commands} conditioning each token both on the image
features and on the previously generated tokens. We also use a
differentiable attention mechanism (Spatial Transformer
Networks:~\cite{jaderberg2015spatial}) to let the model attend to
different regions of the image while predicting each token. For
example, the network first decides to emit the \verb|circle| token
conditioned on the image features, then it emits the $x$ coordinate of
the circle conditioned on the image features and the \verb|circle|
token, and finally it predicts the $y$ coordinate of the circle
conditioned on the image features, the \verb|circle| token, and the
$x$ coordinate. We constrain
coordinates to lie on a discrete $16\times 16$ grid, and later explain
how to remove this restriction.  See supplement for the full details
of the architecture, which we implemented in
Tensorflow~\cite{tensorflow2015-whitepaper}.
%% \remark{There are some more details that are important to provide
%%   about this architecture in the supplement: the
%%   functional form(s) of the probability distributions over tokens, the
%%   network layer sizes, which MLPs share parameters, etc.}

@@

Training does not require backpropagation across the entire sequence
of drawing commands: drawing to the canvas `blocks' the gradients,
effectively offloading memory to an external visual store.  In a
sense, this model is like an autoregressive variant of
Attend-Infer-Repeat~\cite{eslami1603attend}, but critically, \emph{without a learned
  recurrent memory state}, e.g. an RNN.  We can do without an RNN
because we have access to ground truth (image, trace) pairs.  This
allows us to handle scenes with many objects without
worrying about the conditioning of gradients as they propagate
over long sequences, and also makes training more straightforward (it
is just maximum likelihood estimation of the model parameters).

Our network suffices to ``derender'' synthetic images like those shown in
Fig.~\ref{trainingData}.  We can perform a beam search decoding to
recover what the network thinks is the most likely trace for
images like these, recovering traces maximizing $\probability_\theta
[T|I]$. But, if the network makes a mistake (predicts an incorrect
line of code), it has no way of recovering from the error.  In order
to derender an image with $n$ objects, it must correctly predict $n$
drawing commands -- so its probability of success will decrease
exponentially in $n$, assuming it has any nonzero chance of making a
mistake.  For added robustness as $n$ becomes large, we treat the
neural network outputs as proposals for a Sequential Monte Carlo (SMC) sampling scheme~\cite{SMCBook}.  For
the SMC sampler, we use pixel-wise distance as a surrogate for a
likelihood function. The SMC sampler is designed to produce samples
from the distribution $\propto L(I|\text{render}(T))
\probability_\theta[T|I]$, where $L(\cdot | \cdot):\text{image}^2\to
\mathbb{R}$ uses the distance between two images as a proxy for a
likelihood. Unconventionally, the target distribution of the SMC sampler
includes the likelihood under the proposal distribution.
Intuitively, both the proposal
distribution and the distance function offer complementary signals for
whether a drawing command is correct.
%Empirically, accuracy is much better when we combine these signals.

Fig.~\ref{syntheticResults}
compares the neural network with SMC against the neural network by
itself or SMC by itself.  Only the combination of the two passes a
critical test of generalization: when trained on images with $\leq 12$
objects, it successfully parses scenes with many more objects than the
training data.
We compare with a natural baseline that models the problem as a kind of image captioning (LSTM in Fig.~\ref{syntheticResults}).
Given the target image, this baseline produces the trace in one shot by
using a CNN to extract features of the input which are passed to an LSTM which finally predicts
the trace token-by-token.
This architecture is used in several successful neural models of image captioning (e.g.,~\cite{vinyals2015show}),
but, for this domain, does not suffice to parse cluttered scenes with many objects.

@@

\subsection{Generalizing to hand drawings}
A practical application of our neural network is the automatic conversion of hand drawings into a subset of \LaTeX.
 We train the model
to generalize to hand drawings by introducing noise into the
renderings of the training target images.
We designed this noise process to introduce the kinds of variations found in hand drawings (Fig.~\ref{handDrawingExamples}; see supplement for details).

@@

Our neurally-guided SMC procedure
used pixel-wise distance as a surrogate for a likelihood function ($L(\cdot|\cdot)$ in section~\ref{neuralNetworkSection}).
But pixel-wise distance fares poorly on hand drawings, which never exactly match
the model's renders.
So, for hand drawings,
we \emph{learn} a surrogate likelihood function,
$L_{\text{learned}}(\cdot|\cdot)$.
The density $L_{\text{learned}}(\cdot|\cdot)$ is predicted by a convolutional network that we train to predict
the distance between two traces conditioned upon their renderings.
We train our likelihood surrogate to approximate the symmetric difference,
which is  the number of drawing commands by which two traces differ:
\begin{multline}
-\log L_{\text{learned}}(\text{render}(T_1)|\text{render}(T_2))\approx\\ |T_1 - T_2| + |T_2 - T_1|\label{symmetricDistance}
\end{multline}
%Intuitively,
%$L_{\text{learned}}(\cdot|\cdot)$ approximates the distance between
%the trace we want and the trace we have so far.
Pixel-wise distance
metrics are sensitive to the details of how
arrows, dashes, and corners are drawn -- but we wish to be invariant
to these details.
So, we learn a distance metric over images that
approximates the distance metric in the search space over traces.

We drew 100 figures by hand; see Fig.~\ref{lotsOfHandDrawings}.
These were drawn carefully but not perfectly.
Because our model assumes that objects are snapped to a $16\times 16$ grid, 
we made the drawings on graph paper.

For each drawing we annotated a ground truth trace and asked the neurally guided SMC sampler
to produce many candidate traces for each drawing.
For 63\% of the drawings, the Top-1 most likely sample exactly matches the
ground truth; with more samples, the model finds traces
that are closer to the ground truth annotation (Fig.~\ref{drawingIntersectionOverUnion}).
Because the model sometimes makes mistakes % on hand drawings,
we envision it working as follows:
a user sketches a diagram,
and the system responds by proposing a few candidate interpretations.
The user could then select the one closest to their intention and edit it if necessary.

@@

We constrain the space of allowed programs by writing down a context
free grammar over a space of programs. Although it might be desirable
to synthesize programs in a Turing-complete language such as Lisp or
Python, a more tractable approach is to specify what in the program
languages community is called a Domain Specific Language (DSL)~\cite{polozov2015flashmeta}. Our DSL (Tbl.~\ref{DSL})
encodes prior knowledge of what graphics programs tend to look like.

@@

Given the DSL and a trace $T$, we want to recover a program that both evaluates to $T$
and, at the same time, is the ``best'' explanation of $T$.
For example, we might prefer more general programs or, in the spirit of Occam's razor,
prefer shorter programs.
We wrap these intuitions up into a cost function over programs,
and seek the minimum cost program consistent with $T$:
\begin{equation}
  \text{program}(T) = \argmin_{p\in \text{DSL, s.t. }p \text{ evaluates to } T} \text{cost}(p)\label{programObjective}
\end{equation}
We define the
cost of a program to be the number of Statement's it contains (Tbl.~\ref{DSL}).
We also penalize using many different numerical constants; see supplement.
%\remark{The flow here is a bit off/backwards. ``We want a program that evaluates to $T$ and also minimizes some measure of program cost''---why do we care about cost? It'd be better to start by making a ``Bayesian Occam's razor'' appeal (e.g. the most compact/general program is the more likely explanation) and then saying that one way to do this is to minimize a cost function which is proportional to program length.}

The constrained optimization problem in
Eq.~\ref{programObjective} is intractable in general, but there
exist efficient-in-practice tools for finding exact solutions to such
program synthesis problems. We use the state-of-the-art Sketch
tool~\cite{solar2008program}.
%Describing Sketch's program synthesis
%algorithm is beyond the scope of this paper; see~\cite{solar2008program}.
At a
high level, Sketch takes as input a space of programs, along with a
specification of the program's behavior and optionally a cost
function.  It translates the synthesis problem into a constraint
satisfaction problem and then uses a SAT solver to find a
minimum-cost program satisfying the specification.  In exchange for
having no guarantee on the time required to find a minimum
cost solution, it comes with the guarantee that it will always find a
globally optimal program.
Synthesis times vary from minutes to hours,
with 27\% of the drawings timing out the synthesizer after 1 hour.
See Fig.~\ref{exampleSynthesisResults} for examples of the kinds of programs our system recovers.
A main impediment to our use of these general techniques is
the prohibitively high cost of searching for programs.
We next describe how to learn to synthesize programs much faster (Section~\ref{learningASearchPolicy}),
timing out on 2\% of the drawings and solving most problems in under a minute.

\begin{figure}
  \vspace{-0.6cm}
\centering  \begin{tabular}{m{1.5cm}llc}
%    \toprule
    \textbf{Drawing}&\textbf{Trace}&\textbf{Program}&%\begin{tabular}{c}
      \textbf{Compression factor}%\\\textbf{factor}
      %      \end{tabular}\\
      \\
    \midrule
    \includegraphics[width = \exampleDrawingSize]{figures/expert-29-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Line(2,15, 4,15)
Line(4,9, 4,13)
Line(3,11, 3,14)
Line(2,13, 2,15)
Line(3,14, 6,14)
Line(4,13, 8,13)
\end{lstlisting}
\end{minipage}&     \begin{minipage}{\exampleProgramSize} \begin{lstlisting}
for(i<3)
 line(i,-1*i+6,
      2*i+2,-1*i+6)
 line(i,-2*i+4,i,-1*i+6)
       \end{lstlisting}
     \end{minipage}&$\frac{6}{3} = 2\text{x}$\\\midrule
     \includegraphics[width = \exampleDrawingSize]{figures/expert-52-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Line(5,13, 2,10),
  arrow = True)
Circle(5,9)
Circle(8,5)
Line(2,8, 2,6), 
  arrow = True)
Circle(2,5)
Circle(5,14)
Circle(2,9)
\end{lstlisting}
\emph{... etc. ...; 13 lines}
\end{minipage}&
             \begin{minipage}{\exampleProgramSize}\begin{lstlisting}
circle(4,10)
for(i<3)
 circle(-3*i+7,5)
 circle(-3*i+7,1)
 line(-3*i+7,4,-3*i+7,2,
      arrow=True)
 line(4,9,-3*i+7,6,
      arrow=True)
\end{lstlisting}
\end{minipage}&$\frac{13}{6} = 2.2\text{x}$\\\midrule
    \includegraphics[width = \exampleDrawingSize]{figures/expert-38-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Circle(5,8)
Circle(2,8)
Circle(8,11)
Line(2,9, 2,10)
Circle(8,8)
Line(3,8, 4,8)
Line(3,11, 4,11)
Line(8,9, 8,10)
Circle(5,14)
\end{lstlisting}
  \emph{... etc. ...; 21 lines}
\end{minipage}&\begin{minipage}{\exampleProgramSize}
\begin{lstlisting}
for(i<3)
 for(j<3)
  if(j>0)
   line(-3*j+8,-3*i+7,
        -3*j+9,-3*i+7)
  line(-3*i+7,-3*j+8,
       -3*i+7,-3*j+9)
  circle(-3*j+7,-3*i+7)
\end{lstlisting}
\end{minipage}&$\frac{21}{6} = 3.5\text{x}$\\\midrule
\includegraphics[width = \exampleDrawingSize]{figures/expert-75-extra-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Line(11,14, 13,14), 
  arrow = True)
Circle(10,10)
Line(10,13, 10,11), 
  arrow = True)
Line(7,14, 9,14), 
  arrow = True)
Circle(6,10)
Circle(2,10)
\end{lstlisting}
    \emph{... etc. ...; 15 lines}
\end{minipage}&
\begin{minipage}{\exampleProgramSize}
    \begin{lstlisting}
for(i<4)
 line(-4*i+13,4,-4*i+13,2,
      arrow=True)
 for(j<3)
  if(j>0)
   circle(-4*i+13,4*j+-3)
  line(-4*j+10,5,-4*j+12,5,
       arrow=True)
\end{lstlisting}
  \end{minipage}&$\frac{15}{6} = 2.5\text{x}$\\\midrule

\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Circle(2,8)
Rectangle(6,9, 7,10)
Circle(8,8)
Rectangle(6,12, 7,13)
Rectangle(3,9, 4,10)
Circle(2,14)
Circle(8,14)
Circle(5,11)
Rectangle(3,12, 4,13)
\end{lstlisting}
  \end{minipage}&\begin{minipage}{\exampleProgramSize}
\begin{lstlisting}
reflect(y=8)
 for(i<3)
  if(i>0)
   rectangle(3*i-1,2,3*i,3)
  circle(3*i+1,3*i+1)
\end{lstlisting}
\end{minipage}&$\frac{9}{5} = 1.8\text{x}$ %\\\bottomrule
  \end{tabular}
\caption{Example drawings (left), their ground truth traces (middle left), and programs synthesized from these traces (middle right). Compared to the trace the programs are more compressive (right: programs have fewer lines than traces) and automatically group together related drawing commands. Note the nested loops in the Ising model, special case conditionals for the HMM, combination of symmetry and iteration in the bottom figure, and affine transformations in the top figure and second figure to bottom.}\label{exampleSynthesisResults}
  \vspace{-0.7cm}
  \end{figure}

@@

\subsection{Related work}
